WORCS: A Workflow for Open Reproducible Code in Science

### 2020-11-17 paper/Manuscript_DS.Rmd
Further discuss GitHub releases

An integral part of WORCS is the *explicit* version control software Git
(www.git-scm.com).
Git tracks changes to files on a line-by-line basis.
The user can store these changes by making a "commit": a 
snapshot of the version controlled files.
Each "commit" can contain as many changes as desired;
for example, a whole new paragraph, or many small changes made to address a single reviewer comment.
Each commit is further tagged with a message, describing the changes.
The `R` implementation of `worcs` contains a convenience function,
`git_update("your commit message")`,
which creates a new commit for all changes since the previous commit,
labels it with "your commit message", and pushes these changes to the remote repository.
It is, effectively, a Git "quick save" command.

From an open science perspective, Git is particularly appealing because it
retains a complete historical backlog of all commits.
This can be used, for example, to verify that authors executed the steps outlined in a preregistration.
Users can also compare changes between different commits,
or go back to a previous version of the code (for example,
after making a mistake, or to replicate a previous version of the results).
A project version controlled with Git is called a "Repository", or Repo.
Git only version controls files explicitly committed by the user. Moreover, it
is possible to prevent files from being version controlled - which is useful for
privacy sensitive data.
A `.gitignore` file lists files that should not be version-controlled.
These files thus exist only on the user's computer.

The functionality of Git is enhanced by services such as GitHub
(https://github.com). GitHub is best understood as a cloud storage service
with social networking functionality. The cloud storage aspect
of GitHub allows you to "clone" (copy) a local Git repository to
the GitHub website, as a backup or research archive. The social network aspect
comes into play when a repository on GitHub is set to "public": This allows 
other researchers to peruse the repository and see how the work was done; clone
it to their own computer to replicate the original work or apply the methods to
their own data; open "Issues" to ask questions or give feedback on the project,
or even send a "Pull request" with suggested changes to the text or code for
your consideration. Git and GitHub shine as tools for collaboration, because
different people can simultaneously work on different parts of a project, and
their changes can be compared and automatically merged on the website. Even on
solo projects, working with Git/GitHub has many benefits: Staying organized,
being able to start a new study with a clone of an old, similar repository, or
splitting off an "experimental branch" to try something new, while retaining the
ability to "revert" (return) to a previous state of the project, or to "merge"
(incorporate) the experimental branch.

>>>>

An integral part of WORCS is the *explicit* version control software Git
(www.git-scm.com).
A project version controlled with Git is called a "repository", or repo.
Git tracks changes to files in the repository on a line-by-line basis.
The user can store these changes by making a "commit": a 
snapshot of the version controlled files.
Each "commit" can contain as many changes as desired;
for example, a whole new paragraph, or many small changes made to address a single reviewer comment.
Each commit is further tagged with a message, describing the changes.
The `R` implementation of `worcs` contains a convenience function,
`git_update("your commit message")`,
which creates a new commit for all changes since the previous commit,
labels it with "your commit message", and pushes these changes to the remote repository.
It is, effectively, a Git "quick save" command.

From an open science perspective, Git is particularly appealing because it
retains a complete historical backlog of all commits.
This can be used, for example, to verify that authors executed the steps outlined in a preregistration.
Users can also compare changes between different commits,
or go back to a previous version of the code (for example,
after making a mistake, or to replicate a previous version of the results).
Git only version controls files explicitly committed by the user. Moreover, it
is possible to prevent files from being version controlled - which is useful for
privacy sensitive data.
A `.gitignore` file lists files that should not be version-controlled.
These files thus exist only on the user's computer.

The functionality of Git is enhanced by services such as GitHub
(https://github.com). GitHub is best understood as a cloud storage service
with social networking functionality. The cloud storage aspect
of GitHub allows you to "clone" (copy) a local Git repository to
the GitHub website, as a backup or research archive.
The GitHub platform offers additional functionality over Git. 
One such feature is that specific stages in the lifecycle of a project
can be [tagged as a "release"](https://help.github.com/en/github/administering-a-repository/creating-releases):
A named downloadable snapshot of a specific state of the project,
that is prominently featured on the GitHub project page.
The WORCS procedure encourages users to create releases to demarcate project milestones
such as preregistration and submission to a journal.
Releases are also useful to mark the submission of a revised manuscript, and publication.

The social network aspect of GitHub
comes into play when a repository is made "public": This allows 
other researchers to peruse the repository and see how the work was done; clone
it to their own computer to replicate the original work or apply the methods to
their own data; open "Issues" to ask questions or give feedback on the project,
or even send a "Pull request" with suggested changes to the text or code for
your consideration. Git and GitHub shine as tools for collaboration, because
different people can simultaneously work on different parts of a project, and
their changes can be compared and automatically merged on the website. Even on
solo projects, working with Git/GitHub has many benefits: Staying organized,
being able to start a new study with a clone of an old, similar repository, or
splitting off an "experimental branch" to try something new, while retaining the
ability to "revert" (return) to a previous state of the project, or to "merge"
(incorporate) the experimental branch.

### 2020-11-17 paper/reviews_ds.Rmd
Further discuss GitHub releases

>>>>

The GitHub function of creating "releases" also helps mark the "course-grained" progress of a research project. In the WORCS procedure, we encourage taggin the preprint and submitted version of a manuscript as a "release" on GitHub. Such releases are prominently featured on the GitHub project page, and are accompanied by a downloadable archive of the historic state of the project.
To further address the Reviewer's comment, we now mention this in the manuscript text as well:

> The GitHub platform offers additional functionality over Git. 
One such feature is that specific stages in the lifecycle of a project
can be [tagged as a "release"](https://help.github.com/en/github/administering-a-repository/creating-releases):
A named downloadable snapshot of a specific state of the project,
that is prominently featured on the GitHub project page.
The WORCS procedure encourages users to create releases to demarcate project milestones
such as preregistration and submission to a journal.
Releases are also useful to mark the submission of a revised manuscript, and publication.
